---
title: "Class 15: Genomics I"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```


# Read sample genotypes data from ENSEMBLE

We downloaded genotype data from ENSEMBLE on the MXL mexican Ancestry in Los Angeles, CA.

What proportion of this data are G/G etc?

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

We want to look at the second column that contains the genotype information
```{r}
table(mxl$Genotype..forward.strand.)
```

## RNA-Seq result analysis for different genotypes of this SNP
```{r}
expr <- read.table("https://bioboot.github.io/bggn213_W20/class-material/rs8067378_ENSG00000172057.6.txt")
head(expr)
table(expr$geno)
```

```{r}
mean(expr$exp[expr$geno=="G/G"])

summary(expr[expr$geno=="G/G",]$exp)

summary(expr[expr$geno=="A/G",]$exp)

summary(expr[expr$geno=="A/A",]$exp)

```

## Box plot

We will use the `boxplot()` function and the input data will be **expr**
```{r}
boxplot(exp ~ geno, data=expr)
```

