---
title: "Structural Bioinformatics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1.PDB Statistics

Download a CSV file from the PDB site (accessible from “Analyze” > “PDB Statistics” > “by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project and use it to answer the following questions:

>Q: What percentage of structures in the PDB are solved by X-Ray and Electron Microscopy.

```{r}
stats<-read.csv("Data Export Summary.csv",row.names = 1)
per<-round(stats$Total/sum(stats$Total)*100,2)
#Alternative: per<-stats[,"Total"]/sum(stats[,"Total"])
names(per)<-rownames(stats)
per
```
A: X-Ray:88.95 Electron Microscopy: 2.72
>Q: What proportion of structures in the PDB are protein?

```{r} 
per.protein<-round(sum(stats$Proteins)/sum(stats$Total)*100,2)
per.protein
```
A: 92.69%

>Q: Type HIV in the PDB website search box on the home page and determine how many HIV-1 protease structures are in the current PDB?

A: 1,298

## 2.Using bio3d to analyse protein structure data

Use the bio3d package from CRAN.

```{r}
library(bio3d)
pdb<-read.pdb("1hsg")
pdb
```

```{r}
attributes(pdb)
```
```{r}
## check installed packages
#   installed.packages()

## install packages from CRAN
#install.packages("ggplot2") 
#install.packages("ggrepel")
#install.packages("devtools")
#install.packages("BiocManager")

## install packages from BioCondo
#BiocManager::install("msa")

## install packages from github
#devtools::install_bitbucket("Grantlab/bio3d-view")
```

Adenylate kinase structures from the PDB for analysis

```{r}
hits <- NULL
hits$pdb.id <- c('1AKE_A','4X8M_A','6S36_A','6RZE_A','4X8H_A','3HPR_A','1E4V_A','5EJE_A','1E4Y_A','3X2S_A','6HAP_A','6HAM_A','4K46_A','4NP6_A','3GMT_A','4PZL_A')
```

Download and parse these structures
```{r}
##Download related PDB files
##Excecute in R Console
#files <- get.pdb(hits$pdb.id, path="pdbs", split=TRUE, gzip=TRUE)
```

```{r}
# Align releated PDBs
pdbs <- pdbaln(files, fit = TRUE, exefile="msa")
```
```{r}
# Vector containing PDB codes for figure axis
ids <- basename.pdb(pdbs$id)

# Draw schematic alignment
plot(pdbs, labels=ids)
```

```{r}
library(bio3d.view)
library(rgl)
```


```{r}
library(rgl)
view(pdbs)
rglwidget(elementId = "plot3drgl")
```

```{r}
pc<-pca(pdbs)
plot(pc)
```

```{r}
plot(pc,1:2)
```

Make a PCA trajectory

```{r}
pc1 <- mktrj(pc, pc=1, file="pc_1.pdb")
```

