---
title: "Rev-Erba in Th17 RNA-seq Practice"
output: github_document
---
#### This is a demo analysis for the RNA-seq data from Amir et al., 2018, Cell Reports 25, 3733–3749.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this work, we use REV-ERBa WT versus KO RNA-seq from Th17 cells [GEO: GSE122726](https://ftp.ncbi.nlm.nih.gov/geo/series/GSE122nnn/GSE122726/suppl/GSE122726_processed_218.xlsx)
There're two biological replicates for each genotypes in this experiment.


#### 1. Load `DESeq2` package for differential gene expresion analysis.
```{r}
library(DESeq2)
library(dplyr)
```


#### 2. DESeq2 Required Inputs
As input, the DESeq2 package expects **(1)** a data.frame of **count data** (as obtained from RNA-seq or another high-throughput sequencing experiment) and **(2)** a second data.frame with information about the samples - often called sample metadata (or `colData` in DESeq2-speak because it supplies metadata/information about the columns of the countData matrix). 

#### 3. Load count data from GEO database
**Note**: The count values must be raw counts of sequencing reads. This is important for **DESeq2**'s statistic model to hold.

```{r}
# load count table 
count<- read.csv("GSE122726 WTVSKO.csv",stringsAsFactors = FALSE,row.names = 1)
rownames(count) <- gsub("\t", "", rownames(count))
head(count)
```

Some genes have **zero** read counts for all 4 samples. We tidy up the counts data to exclude genes with **zero** value.
```{r}
filter.Count <- count[rowSums(count)!=0,]
head(filter.Count)
```



2. Import countData and colData

Create a `meta` for metadata
```{r}
id <- colnames(filter.Count)
```


```{r}
meta <- data.frame("id"=id[1:4], "condition" = c(rep("WT",2),rep("KO",2)), stringsAsFactors = FALSE)
```


```{r}
dds <- DESeqDataSetFromMatrix(countData=filter.Count, 
                              colData=meta, 
                              design=~condition
                              )
#set WT as reference level
dds$condition <- relevel(dds$condition, ref = "WT")

```

```{r}
dds <- DESeq(dds)
head(dds)
```

Get Deseq results

```{r}
res <- results(dds)
summary(res)
```

Subset the DEG with adjust p value lower than 0.05
```{r}
resAdjSig05 <- subset(as.data.frame(res),padj <0.05)
head(resAdjSig05)
```


```{r}
resAdjSig05up <- subset(resAdjSig05,resAdjSig05$log2FoldChange>0)
resAdjSig05down <- subset(resAdjSig05,resAdjSig05$log2FoldChange<0)
head(resAdjSig05up)
head(resAdjSig05down)
```

Save result with DE gene ordered by padj
```{r}
write.csv(resAdjSig05, "adjsignif05_results.csv")
write.csv(resAdjSig05up, "adjsignif05_results_up.csv")
write.csv(resAdjSig05down, "adjsignif05_results_down.csv")

```



##Plot Deseq Result

#MA-plot
```{r}
# unshrunken
plotMA(res,ylim=c(-2,2))
```

```{r}
#Shrunken
res_shrink <- lfcShrink(dds, contrast=c("condition","KO","WT"), res=res)
plotMA(res_shrink,ylim=c(-2,2))
```
```{r}
#plot p-value distribution
hist(resAdjSig05$pvalue)
```

```{r}
# Plot dispersion estimates
plotDispEsts(dds, main="DESeq2")
```


#Plot counts

This is to examine the counts of reads for a single gene across the groups. `plotCounts()` normalizes counts by sequencing depth and adds a pseudocount of 1/2 to allow for log scale plotting. 
```{r}
plotCounts(dds, gene=which.min(res$padj),intgroup = "condition")

```

Blind dispersion estimation
```{r}
rld <- rlog(dds,blind=FALSE)
vsd <- varianceStabilizingTransformation(dds,blind=FALSE)
vsd.fast <- vst(dds,blind=FALSE)
head(assay(rld),3)
head(assay(vsd),3)
head(assay(vsd.fast),3)


```

## Heatmap of the count matrix

```{r}
library("pheatmap")
select <- order(rowMeans(counts(dds,normalized=TRUE)),decreasing =TRUE)[1:20]
nt <- normTransform(dds) #defaults to log2(x+1)
log2.norm.counts <- assay(nt)[select,]
df <- data.frame("condition" = c(rep("WT",2),rep("KO",2)))
rownames(df) <- meta$id
pheatmap(log2.norm.counts,cluster_rows = FALSE,show_rownames = FALSE,cluster_cols = FALSE,annotation_col =df)
pheatmap(assay(rld)[select,],cluster_rows = FALSE,show_rownames = FALSE,cluster_cols = FALSE,annotation_col =df)
pheatmap(assay(vsd)[select,],cluster_rows = FALSE,show_rownames = FALSE,cluster_cols = FALSE,annotation_col =df)

```

??? Not sure how to do it
```{r}
select.padj <- order(res$padj,res$log2FoldChange)[1:50]

df <- data.frame("condition" = c(rep("WT",2),rep("KO",2)))
rownames(df) <- meta$id

pheatmap(assay(vsd)[select.padj,],cluster_rows = FALSE,show_rownames = FALSE,cluster_cols = FALSE,annotation_col =df)

```

```{r}
sampleDists <- dist(t(assay(rld)))
```

```{r}
library("RColorBrewer")
sampleDistMatrix <- as.matrix(sampleDists)
rownames(sampleDistMatrix) <- paste(rld$condition,rld$id)
colnames(sampleDistMatrix) <- NULL
colors <- colorRampPalette(rev(brewer.pal(9,"Blues")))(100)
pheatmap(sampleDistMatrix,
         clustering_distance_rows = sampleDists,
         clustering_distance_cols = sampleDists,
         col=colors)
```

PCA plot
```{r}
plotPCA(rld,intgroup = "condition")
```




```{r}
library(ggplot2)
data <- plotPCA(rld, intgroup="condition", returnData=TRUE)
percentVar <- round(100 * attr(data, "percentVar"))
ggplot(data, aes(PC1, PC2, color=condition)) +
    geom_point(size=3) +
    xlab(paste0("PC1: ",percentVar[1],"% variance")) 
    ylab(paste0("PC2: ",percentVar[2],"% variance"))
```


```{r}
library(org.Mm.eg.db)
library("AnnotationDbi")
```

```{r}
columns(org.Mm.eg.db)
```
```{r}
res$entrez <-unname(mapIds(org.Mm.eg.db, rownames(res), "ENTREZID", "SYMBOL"))
res$ensembl <- unname(mapIds(org.Mm.eg.db, rownames(res), "ENSEMBL", "SYMBOL"))
```
```{r}
head(res)
```
```{r}
resOrdered <- res[order(res$pvalue),]
head(resOrdered)
```

```{r}
write.csv(as.data.frame(resOrdered), file="results.csv")
```

heatmap for gene clustering
```{r}
library("genefilter")
topVarGenes <- head(order(rowVars(assay(rld)), decreasing = TRUE), 100)

```

```{r}
library("gplots")
```


```{r}
#head(rownames(resOrdered))
LFCvsdup <- subset(assay(vsd), rownames(assay(vsd)) %in% rownames(resAdjSig05up))
LFCvsddown <- subset(assay(vsd), rownames(assay(vsd)) %in% rownames(resAdjSig05down))


LFCtopVarGenesup <- head(order(rowVars(LFCvsdup),decreasing=TRUE),40)
LFCtopVarGenesdown <- head(order(rowVars(LFCvsddown), decreasing = TRUE),67)

#Heatmap for topVarGenes from resAdjSig05up
heatmap.2(LFCvsdup[LFCtopVarGenesup,], scale = "row", trace = "none", dendrogram = "column",
col = colorRampPalette(rev(brewer.pal(9, "RdBu")))(255))

heatmap.2(LFCvsddown[LFCtopVarGenesdown,], scale = "row", trace = "none", dendrogram = "column",Colv=FALSE,
col = colorRampPalette(rev(brewer.pal(9, "RdBu")))(255),labRow = rownames(LFCvsddown)[LFCtopVarGenesdown])
```

```{r}
pheatmap(LFCvsdup[LFCtopVarGenesup,],scale = "row",clustering_distance_rows = "correlation",cluster_rows = FALSE,show_rownames = TRUE,cluster_cols = FALSE,annotation_col =df)
```



## GSEA with `gage` package
```{r}
library(gage)
library(pathview)

```

```{r}
library(gageData)
data("kegg.sets.mm")
data(sigmet.idx.mm)
```

```{r}
kegg.sets.M=kegg.sets.mm[sigmet.idx.mm]
head(kegg.sets.mm,3)
```

```{r}
foldchanges <-  res$log2FoldChange
names(foldchanges) <- res$entrez
head(foldchanges)
```



```{r}
keggres = gage(foldchanges,gsets = kegg.sets.mm)
attributes(keggres)
```
```{r}
keggl <- keggres$less
keggs <- keggres$stats
```

```{r}
lapply(keggres,head)
```
```{r}
keggrespathways <- data.frame(id=rownames(keggres$greater,keggres$greater %>% tbl_df() %>% filter(row_number()<=5 %>% .$id %>% as.character())))
head(keggrespathways,20)
```


```{r}
keggresids <- substr(keggrespathways$id,start=1,stop=8)
head(keggresids)
```

??
```{r}
#plot_pathway <- function(pid) pathview(gene.data = foldchanges,pathway.id = pid, species = "mm")
#tmp <- sapply(keggresids,function(pid) pathview(gene.data = foldchanges,pathway.id = pid,species="mm"))
```


## GO analysis

```{r}
data("go.sets.mm")
data("go.subs.mm")
gobpsets <- go.sets.mm[go.subs.mm$BP]
gobpres <- gage(foldchanges,gsets = gobpsets,same.dir = TRUE)
lapply(gobpres,head,10)
```

```{r}
gobprespathways <- data.frame(id=rownames(gobpres$greater,gobpres$greater %>% tbl_df() %>% filter(row_number()<=5 %>% .$id %>% as.character())))
head(gobprespathways,20)

```

```{r}
gobpresterm <- substr(gobprespathways$id,12,100)
head(gobpresterm,20)
```



```{r}
#keggrld <- assay(rld)
#head(keggrld)
#geneData(genes = keggresids, exprs = assay(rld)[], outname = outname,txt=T, heatmap = T,limit = 3,scartterplot=TRUE)
```

```{r}
#gobprespathways$id[1]
```


```{r}
#for (gs in gobprespathways$greater){
  #outname=gsub(" |:|/","_",substr(gs,12,100))
  #geneData(genes = kegg.sets.mm[[gs]], exprs = assay(rld), outname=outname,txt=T,heatmap=T,limit = 3,scatterplot = T)
#}

```


```{r}
#assay(rld)
#keggresids
```

