---
title: "Class 17 COVID-19 Homework"
output: github_document
---


## Coronavirus

Here we analyze infection data for the 2019 novel Coronavirus COVID-19 (2019-nCoV) epidemic. The raw data is pulled from the Johns Hopkins University Center for Systems Science and Engineering (JHU CCSE) Coronavirus repository.

A CSV file is available here https://github.com/RamiKrispin/coronavirus-csv


```{r}
url <- "https://tinyurl.com/COVID-2019"
virus <- read.csv(url)
tail(virus)
```

> Q1. How many total infected cases are there around the world?

```{r}
sum(virus$cases)
```

```{r}
table(virus$type)
```



> Q2. How many deaths linke to infected cases have there been?

```{r}
inds<- virus$type=="death"
sum(virus$cases[inds]) # sum(virus[inds,"case"])
```

> Q3. What is the overall death rate?

```{r}
sum(virus$case[inds])/sum(virus$case)*100

```

> Q4. What is the death rate in "Mainland China"?

```{r}
country <- virus$Country.Region=="Mainland China"

total_case <- sum(virus$cases[country])

country_death <- virus$type=="death" & virus$Country.Region=="Mainland China"

total_death <- sum(virus$cases[country_death])

death_rate <- total_death/total_case*100

round(death_rate,2)
```

> Q5. What is the death rate in Italy, Iran and the US?

```{r}
# Define a function to calculate death rate for countries

covidDR<-function(x){country <- virus$Country.Region==x

total_case <- sum(virus$cases[country])

country_death <- virus$type=="death" & virus$Country.Region==x

total_death <- sum(virus$cases[country_death])

death_rate <- total_death/total_case*100

round(death_rate,2)
}
```

```{r}
# Calculate the death rates for Italy, Iran and the US
covidDR("Italy")
covidDR("Iran")
covidDR("US")
```