---
title: 'Class 10: Machine Learning Project'
author: "Jiayi Dong"
date: "2/7/2020"
output: github_document
---

## The goal of this session is to explore a complete analysis using the unsupervised learning techniques covered in the last class. We'll extend what we've learned by combining PCA as a preprocessing step to clustering using data that consist of measurements of cell nuclei of human breast masses. 

The data itself comes from the Wisconsin Breast Cancer Diagnostic Data Set first reported by K. P. Benne and O. L. Mangasarian: "Robust Linear Programming Discrimination of Two Linearly Inseparable Sets".

```{r}
wisCan.df<- read.csv("https://bioboot.github.io/bimm143_S18/class-material/WisconsinCancer.csv")
```


```{r}
# select column 3 to 32

wisCan.data<-as.matrix(wisCan.df[,3:32])

```

>Q. How many patients are there in the dataset?

```{r}
nrow(wisCan.df)
```

>Q. How many cancer and non-cancer patients are there?

```{r}
table(wisCan.df$diagnosis)
# sum(wisCan.df$diagnosis=="B")
```


```{r}
grep("mean",colnames(wisCan.df), value=TRUE)
```


```{r}
pca <- prcomp(wisCan.data,scale=TRUE)
summary(pca)
```

```{r}
hc <- hclust(dist(wisCan.data))
hc.cut <- cutree(hc,k=3)
plot(hc)
hc
```

```{r}
pca.hc <- hclust(dist(pca$x[,1:3]),method = "ward.D2")
grps <- cutree(pca.hc, k=2)
table(grps)
plot(pca$x[,1:2], col=grps)

```

```{r}
#url <- "new_samples.csv"
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(pca, newdata=new)
npc
```

```{r}
plot(pca$x[,1],pca$x[,2],col=wisCan.df$diagnosis)
plot(npc[,1],npc[,2],col="blue")
```


